---
title: "QMM- Assignment 6"
author: "Sai Gautham Sabhavathu"
date: "2022-11-20"
output: pdf_document
---

```{r}
library(lpSolve)
library(lpSolveAPI)


AP_HUB <- read.lp("C:/Users/gauth/OneDrive/Desktop/AP_HUB.lp")
AP_HUB
```

The table below provides 
an estimate of the number of workers needed each day of the week. 
```{R}
DAYS_AND_WORKERS <- matrix(c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday", 18,27,22,26,25,21,19), ncol = 2, byrow = F)

colnames(DAYS_AND_WORKERS) <- c("Days_Of_The_Week","Workers_Required")
as.table(DAYS_AND_WORKERS)
```

Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an additional $25 per day. The possible shifts and salaries for package handlers are: 

```{r}
Shift_DaysOff_Wage <- matrix(c(1,2,3,4,5,6,7,
                               "Sunday and Monday","Monday and Tuesday","Tuesday and Wednesday","Wednesday and Thursday","Thursday and Friday","Friday and Saturday","Saturday and Sunday",
                               "$775","$800","$800","$800","$800","$775","$750"), ncol = 3, byrow = F)

colnames(Shift_DaysOff_Wage) <- c("Shift", "Days_Off", "Wage")
as.table(Shift_DaysOff_Wage)
```

```{r}
solve(AP_HUB)
```

```{r}
get.objective(AP_HUB)

```
The total cost is $25675.
```{r}
get.variables(AP_HUB)
```

From the above, we can derive :-
x1 = 2, Which means 2 workers have been assigned to Shift 1.
x2 = 4, Which means 4 workers have been assigned to Shift 2.
x3 = 5, Which means 5 workers have been assigned to Shift 3.
x4 = 0, Which means 0 workers have been assigned to Shift 4.
x5 = 8, Which means 8 workers have been assigned to Shift 5.
x6 = 1, Which means 1 worker has been assigned to Shift 6.
x7 = 13, Which means 13 workers have been assigned to Shift 7.

Hence, the workers availabe for each day is 
```{r}
Workers_Available <- matrix(c(0,4,5,0,8,1,0,0,0,5,0,8,1,13,2,0,0,0,8,1,13,2,4,0,0,8,1,13,2,4,5,0,0,1,13,2,3,4,0,0,0,13,2,4,5,0,8,0,0), ncol=7,byrow=TRUE)
colnames(Workers_Available)<- c("Shift1", "Shift2", "Shift3", "Shift4", "Shift5", "Shift6", "Shift7")
row.names(Workers_Available) <- c('Sunday', 'Monday', 'Tuesday','Wednesda','Thursday','Friday','Saturday')
Workers_Available

rowSums(Workers_Available)
```